---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r}
library("ggplot2")
```
## Loading and preprocessing the data
```{r}
loaded_data<-read.csv("activity.csv")
```


## What is mean total number of steps taken per day?
```{r}
totals_per_day<-aggregate(loaded_data$steps, list(Day=loaded_data$date), sum)
mean_per_day<-aggregate(loaded_data$steps, list(Day=loaded_data$date), mean)
```
## Mean and median number of steps taken each day
```{r}
hist(totals_per_day$x)
summary(totals_per_day)
```
## What is the average daily activity pattern?
```{r}
avg_per_interval<-aggregate(loaded_data$steps, list(interval=loaded_data$interval), mean, na.rm=TRUE)
ggplot(avg_per_interval, aes(x=interval,y=x))+geom_line()
```

### Finding the Max interval
```{r}
max_interval_X<-subset(avg_per_interval,x==max(avg_per_interval$x))$interval
max_interval_Y<-subset(avg_per_interval,x==max(avg_per_interval$x))$x

ggplot(avg_per_interval, aes(x=interval,y=x))+geom_line()+geom_vline(xintercept = max_interval_X,colour="red")+geom_text(aes(x=max_interval_X, label="Max interval", y=max_interval_Y), colour="red", hjust = -.1)
```

The max average interval is `r max_interval_X` and the max number steps is `r max_interval_Y`


## Imputing missing values
```{r}
numberOfMissingValues<-sum(is.na (loaded_data$steps))

```

The number of missing values is: `r numberOfMissingValues`

Replacing the missing data by the average of taken from similar intervals and 
stroring that in a new data frame:
```{r}
fixed_data <- loaded_data

for (row in 1:nrow(loaded_data)) 
{ 
        if(is.na(loaded_data[row,]$steps)) 
        { 
                fixed_data[row,]$steps = subset(avg_per_interval,interval == loaded_data[row,]$interval)$x }}
```



## Histogram of the total number of steps taken each day after missing values are imputed
```{r}
totals_per_day_imputed<-aggregate(fixed_data$steps, list(Day=fixed_data$date), sum)
mean_per_day_imputed<-aggregate(fixed_data$steps, list(Day=fixed_data$date), mean)

hist(totals_per_day_imputed$x)
summary(totals_per_day_imputed)
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
fixed_data$daytype<- weekdays(as.Date(fixed_data$date))
fixed_data$daytype[fixed_data$daytype %in% c('Saturday', 'Sunday')] <- "weekend"
fixed_data$daytype[fixed_data$daytype != "weekend"] <- "weekday"

fixed_data$daytype<-as.factor(fixed_data$daytype)

avg_fixed_data <- aggregate(steps ~ interval + daytype ,fixed_data,mean)


ggplot(avg_fixed_data, aes(x=interval , y= steps)) + geom_line() + facet_grid(daytype~.)

```






